Letter to Reviewers of Submission \#341

### 2015-06-28 interviews.tex
"informally" interviewed
spellchecker pass and full revision letter

\paragraph*{Web is limiting} Although most analysts appreciated RCloud's sharing features,
it was not a popular tool for exploratory data analysis.
All analysts with R experience prefered to do analysis in another
tool, then paste code into RCloud for sharing, because the web interface got in the
way of their work.

>>>>

\paragraph*{Web is limiting} Although most analysts appreciated RCloud's sharing features,
it was not a popular tool for exploratory data analysis.
All analysts with R experience preferred to do analysis in another
tool, then paste code into RCloud for sharing, because the web interface got in the
way of their work.

### 2015-06-28 relatedwork.tex
spellchecker pass and full revision letter

Heer and Agarwala identify many design considerations for
collaborative visual analytics~\cite{Heer:2008:DCF} that
influenced our work.
RCloud notebooks, and the integrated version control system for them,
described in Section~\ref{sec:notebooks}, address {\it modularity and
granularity}, and {\it artifact histories}.
\emph{Starring}, the means for signaling interest in notebooks, described in
Section~\ref{sec:starring}, addresses social-psychological incentives,
recommendation, and voting and ranking. RCloud's integrated deployment
mechanism, described in Section~\ref{sec:deployment}, addresses the cost of
integration, content export, presentation and view sharing.

@@

There has been noteworthy work on specific techniques
to support collaborative or social code development and data analysis,
such as social bookmarking \cite{Millen:2006:DSB} \cite{Heer:2007:VAV}
and crowdsourcing \cite{Fast:2014:ECS}.
Similarly, there are computational methods to support high
performance execution in incremental code development
environments \cite{Guo:2010:TPI}.
The goal of RCloud is to define an environment in which many such
techniques may be integrated and made available to a broad community.
\begin{figure*}
  \vspace{-1.5em}
\centering
\includegraphics[width=.95\linewidth]{fig/notebook/notebook.pdf}
  \vspace{-1.5em}
\caption{\label{fig:notebook}An RCloud notebook is a sequence of
\emph{cells}, each a snippet of source in one of the supported languages (typically R, but Python and others are also supported) or Markdown. The main creation workflow involves editing notebooks, which are transparently stored as git repositories in GitHub, providing us with easy access to primitives for version tracking. Notebooks can be executed as they're edited (left), or in a standalone viewer (right), via a slightly different URL. This provides a lightweight, low-friction mechanism for sharing results which we discuss in detail in Section~\ref{sec:interviews}. }
  \vspace{-1.5em}
\end{figure*}

One high-level goal is to reduce the gap between implementers and
deployers in visual analytics. The fusion of development with
production operations in software release management (``DevOps''
\cite{Httermann:2012:DD} or ``continuous integration''
\cite{Fowler:2006:Continuous}) is a trend in web services and similar
fields.
%% By making it convenient for data scientists to expend just a
%% little additional effort when creating experiments, we may be able to
%% eliminate the need for programming teams to recreate their work to
%% deploy it in production, which has a high cost in time, expense and
%% accuracy.
%gw{sentence was just too long}
When programming teams recreate the work of data scientists to
deploy it in production, it has a high cost in time, expense and
accuracy. By making it convenient for data scientists to productionize
experiments, we may be able to eliminate the need for this costly step.

>>>>

Heer and Agrawala identify many design considerations for
collaborative visual analytics~\cite{Heer:2008:DCF} that
influenced our work.
RCloud notebooks, and the integrated version control system for them,
described in Section~\ref{sec:notebooks}, address {\it modularity and
granularity}, and {\it artifact histories}.
\emph{Starring}, the means for signaling interest in notebooks, described in
Section~\ref{sec:starring}, addresses social-psychological incentives,
recommendation, and voting and ranking. RCloud's integrated deployment
mechanism, described in Section~\ref{sec:deployment}, addresses the cost of
integration, content export, presentation and view sharing.

@@

There has been noteworthy work on specific techniques
to support collaborative or social code development and data analysis,
such as social bookmarking \cite{Millen:2006:DSB} \cite{Heer:2007:VAV}
and crowdsourcing \cite{Fast:2014:ECS}.
Similarly, there are computational methods to support high
performance execution in incremental code development
environments \cite{Guo:2010:TPI}.
The goal of RCloud is to define an environment in which many such
techniques may be integrated and made available to a broad community.
\begin{figure*}
\centering
\includegraphics[width=.95\linewidth]{fig/notebook/notebook.pdf}
\caption{\label{fig:notebook}An RCloud notebook is a sequence of
\emph{cells}, each a snippet of source in one of the supported languages (typically R, but Python and others are also supported) or Markdown. The main creation workflow involves editing notebooks, which are transparently stored as git repositories in GitHub, providing us with easy access to primitives for version tracking. Notebooks can be executed as they're edited (left), or in a standalone viewer (right), via a slightly different URL. This provides a lightweight, low-friction mechanism for sharing results which we discuss in detail in Section~\ref{sec:interviews}. }
\end{figure*}

One high-level goal is to reduce the gap between implementers and
deployers in visual analytics. The fusion of development with
production operations in software release management (``DevOps''
\cite{Httermann:2012:DD} or ``continuous integration''
\cite{Fowler:2006:Continuous}) is a trend in web services and similar
fields.
%% By making it convenient for data scientists to expend just a
%% little additional effort when creating experiments, we may be able to
%% eliminate the need for programming teams to recreate their work to
%% deploy it in production, which has a high cost in time, expense and
%% accuracy.
%gw{sentence was just too long}
When programming teams recreate the work of data scientists to
deploy it in production, there is a high cost in time, expense and
accuracy. By making it convenient for data scientists to take
experiments into production, we may be able to eliminate the need for this costly step.

### 2015-06-28 revision_letter.tex
spellchecker pass and full revision letter

\section*{Specific changes}

@@

\paragraph*{Comparisons} We have added the comparison to ManyEyes to Table 1.

\paragraph*{Captions} We have added appropriately-sized captions to Figures 2 and 4.

>>>>

\section*{Specific changes, by review criteria}

We will answer the primary reviewer, point by point.

\paragraph*{1. Strengths and weaknesses}
We have kept most of the content of the interviews, but structured it
differently to provide more context, as suggested by a reviewer. On
the other hand, we have chosen to keep the details about the version
control and execution, because they are central to the system design
of RCloud, and they enable the collaboration modes we wish the system
support. As we've argued above, inadequate infrastructure for sharing
and searching for different analyses and visualizations is becoming
the bottleneck in large organizations, and we believe our
infrastructure offers a novel attack on the problem.

\paragraph*{2. Is the work novel, incremental, or previously published? 3. Is this work relevant for VAST?}
We accept that there are no novel visual encodings in RCloud; our goal
was to provide supporting infrastructure for visual and exploratory
analysis as a central tool for discovery in a larger organization. We
believe there are many previously published VAST papers falling into
this broader category, and would like to argue for our submission to
be considered similarly.

\paragraph*{4. Is related work adequately referenced?}
Even though the primary reviewer did not request any changes on this
point, later in the review (specifically in point 7) they mentioned a
missing comparison to ManyEyes in Table 1. We have added this
comparison.

\paragraph*{5. Are the technical results sound?}
We apologize for the lack of other examples, but we chose LDAVis as
one application to showcase RCloud because, to put it completely
frankly, we are not allowed to disclose the other applications being
developed at AT\&T Labs. LDAVis has been given publication clearance
\emph{and} is developed in RCloud. We understand that this is not a
completely satisfactory solution, but again, we believe the
fundamental contribution of RCloud is to reduce the barriers for
exploratory visual analysis in larger organizations, rather than any
specific visual encodings. Since RCloud is built on top of R and
HTML5, any technologies spanning this subspace are usable: LDAVis uses
D3, but one can similarly develop visual encodings using other
techniques such as WebGL.

\paragraph*{6. Is the exposition clear? 7. Should anything be deleted or condensed?}
We have reduced the length of the introduction, as requested, but kept
the system description to most of its current length. We believe that
exposition to be important for the completeness of the manuscript.

\paragraph*{8. Are the figures informative?}
We have extended the captions of the requested figures.

\section*{Other issues}

@@

\paragraph*{Should anything be deleted or condensed?}
We have added the comparison to ManyEyes in Table 1, as suggested by the review.
